---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bqschol 

<!-- badges: start -->
<!-- badges: end -->

The goal of bqschol is to provide an interface to SUB Göttingen's 
big scholarly datasets stored on Google Big Query.

This package is for internal use.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("njahn82/bqschol")
```
## Initialize the connection

Connect to dataset with Crossref metadata snapshots

```{r, message=FALSE}
library(bqschol)
my_con <- bqschol::bgschol_con(
  dataset = "cr_history")
```

Ideally you have a service account token stored as in a json 
to make use of this package. If not available, your Google account
credentials will be requested via the web browser. 

## Table functions

The package provides wrapper for the most common table operations

* `bgschol_list()`: List tables
* `bgschol_tbl()`: Access tables with [https://dplyr.tidyverse.org/](dplyr)
* `bgschol_query()`: Perform of a SQL query and retrieve results
* `bgschol_execute()`: Execute a SQL query on the database

Let's start by listing yearly Crossref historic snapshots.

```{r}
bgschol_list(my_con)
```

We can determine the top publisher by type as of April 2018.
Note that we only stored Crossref records published later than
2007.

```{r, message = FALSE}
library(dplyr)
cr_instant_df <- bgschol_tbl(my_con, table = "cr_apr18")
cr_instant_df %>%
    #top publishers
    dplyr::group_by(publisher) %>%
    dplyr::summarise(n = dplyr::n_distinct(doi)) %>%
    dplyr::arrange(desc(n)) 
```



For more complex tasks, we use SQL. 

```{r}
cc_query <- c("SELECT
  publisher,
  COUNT(DISTINCT(DOI)) AS n
FROM
  cr_apr18,
  UNNEST(license) AS license
WHERE
  REGEXP_CONTAINS(license.URL, 'creativecommons')
GROUP BY
  publisher
ORDER BY
  n DESC
LIMIT
  10")
bgschol_query(my_con, cc_query)
```

`bgschol_execute()` is when new tables shall be created or dropped in
Big Query.

